---
title: "Shiny leaflet example"
author: "Matthew Leonawicz"
output: html_document
runtime: shiny
---

```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, eval=TRUE, message=FALSE, warning=FALSE)
lapply(list.files(pattern="ex_0.*.shiny.R"), read_chunk)
library(shiny)
```

Using leaflet in R is fairly simple and there are plently of online resources and examples available already.
See the RStudio [tutorial](https://rstudio.github.io/leaflet/) to get started if you haven't already.
Below I present a series of examples which build upon each other.
The code displays both point data and raster data on leaflet maps with Shiny app integration.
The focus here is to show the utility of event observation on the server side for integrating the behavior of browser elements, namely, the leaflet map.

##
##
### Version 0 - Use observers

The app does not do much in this version, but it provides a clear view of how to use observers in Shiny apps to ensure that various inputs and outputs remain mutually well behaved.
Here I use `observeEvent` three distinct times to tie together the reactive behavior of instances of `selectInput` and `leafletOutput`.
The entire `ui.R` at this point is:

```{r ui00, eval=FALSE}
```

In `server.R` note the three observers. They constitute the entire code block except for the initial, relatively simple call to `renderLeaflet`.
Comments mark the role performed by each call to `observeEvent`, which:

*    Update the `leafletOutput` map when clicked
*    Update the `selectInput` when the map is clicked
*    Update the `leafletOutput` when the `selectInput` changes

```{r server00, eval=FALSE}
```

Care must be taken to ensure each observer does not update when it shouldn't. One can easily trigger another. Infinite recursion is not hard to imagine in this context.
But fortunately it is not too difficult to make the three `observeEvent` calls work well together.
Use the app here to see the interaction between map clicks and the location selection menu. Version one will build upon this interaction.

```{r, echo=FALSE}
shinyAppFile(
  "ex_00_shiny.R",
  options=list(width="100%", height=400)
)
```

##
##
### Version 1 - Another observer

A single addition of a `checkboxInput` for toggling the location `selectInput` menu.
This requires one more observer in `server.R` but its integration is not complicated.
`ui.R` now looks like the following, with a corresponding nested `conditionalInput` statement:

```{r ui01, eval=FALSE}
```

The server side requires only the following addition of an `observe` call. No changes to the three `observeEvent` calls are necessary.

```{r server01observer, eval=FALSE}
```

Showing the group of points on the map is straightforward.
When hiding the group, I also elect to remove the marker for a selected point if one is selected when the checkbox is deselected.

```{r, echo=FALSE}
shinyAppFile(
  "ex_01_shiny.R",
  options=list(width="100%", height=400)
)
```

I do not want a selected location in the map to return as selected when the checkbox for showing map points is rechecked.
This is a matter of preference and may make more sense in the context of my completed.
Similarly, I hide the `selectInput` when the checkbox is unchecked in the same call to `observe`.

##
##
### Version 2 - Raster data

In the final version of the app, the inclusion of rasterized map layers appears a more fundamental component of the app and the points overlay appears a relatively supplemental feature.
When I designed the app, the raster layers in fact came first. I present this toy example in reverse order so that the first three observers could be explored initially with minimal app clutter.

Adding the rasterized map will require more observers, but they are all less opaque than the first three.
`ui.R` now includes a `sliderInput` for selecting a decade (the raster maps are decadal averages of future climate model projections) and a `checkboxInput` for toggling the map legend.

```{r ui02, eval=FALSE}
```

Reactive objects are added to store the current raster layer, its extracted values, and a corresponding color palette in `server.R`

```{r server02obj, eval=FALSE}
```

and the required observers are:

```{r server02obs, eval=FALSE}
```

In the first call to `observe` above, a rasterized map is removed by `removeTiles` (if present) before a new one is added.
The second observer corresponds to the new `checkboxInput`. `clearControls` will remove an existing legend.

The three reactive objects are one line each, there is only the one reactive output, and by far the bulk of the code in `server.R` is for the six observers.
The version two app here allows for essentially independent use of either the raster data or the point data.

```{r, echo=FALSE}
shinyAppFile(
  "ex_02_shiny.R",
  options=list(width="100%", height=500)
)
```

The color palette is generated reactively, specific to the current raster map, and for this data won't appear to change much as you move the slider.
Things are getting a bit cramped in this display window but I will leave it as is.

##
##
### Version 3 - Point data and modals

Returning to the point data, the app employs a modal popup window to display location-specific data when requested by the user for a currently selected map location.
This further helps to separate and create the distinction between the map data and the point data in the app,
which although related, will be used in different ways to highlight different information.

For the modal, use the [shinyBS](https://ebailey78.github.io/shinyBS/index.html) package.

Use `bsModal` in `ui.R`. Standard reactive inputs and outputs can be contained within. The modal is triggered by the referenced `actionButton`.

```{r ui03, eval=FALSE}
```

On the server side, I add the point data (simply hardcoded here) and simple `renderPlot` and `renderDataTable` outputs to be displayed within the modal window.

```{r server03pointdata, eval=FALSE}
```

Select a location and click the action button to which the modal is tied. Note how the rest of the app is grayed out when the modal is active.

```{r, echo=FALSE}
shinyAppFile(
  "ex_03_shiny.R",
  options=list(width="100%", height=500)
)
```

##
#### Side note
In the version at the time this document I found that the modal itself was also grayed out (making it impossible to interact with and close out, leaving the app running but being unable to interact with it).
I don't know exactly where the original conflict came from, but I solved it by adding custom CSS to set the modal backdrop to have a z-index of 1000.
But if you need a similar fix for `bsModal`, I would not expect this to be a perfect fix in all app contexts.

```
.modal-backdrop {
z-index: 1000;
}
```

##
##
### Additional resources

Code for the series of apps shown here can be found on [Github](https://github.com/leonawicz/nwt).
This includes the toy data sets used here so that you can run the app.

See the more advanced [complete app](http://shiny.snap.uaf.edu/nwtapp/) on which this example is based.

For another example, see [my first app using leaflet](http://shiny.snap.uaf.edu/cc4liteFinal/),
which contains thousands of communities across Alaska and western Canada but which does not use rasterized maps.
